---
title: "Business Intelligence Lab Submission Markdown"
author: "<Specify your name here>"
date: "<Specify the date when you submitted the lab>"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

+---------------------------------------------------+---------------------------------------------+-------------------------+
| **Student ID Numbers and Names of Group Members** | \<122147 Melvin Muhati                      | \<Melvin Muhati 122147  |
|                                                   |                                             |                         |
|                                                   | 121774 Kigen Tuwei                          | Kigen Tuwei 121774      |
|                                                   |                                             |                         |
|                                                   | 120578 Irvine Esikote                       | Irvine Esikote 120578   |
|                                                   |                                             |                         |
|                                                   | 135129 Glenn Wamae                          | Glenn Wamae 135129      |
|                                                   |                                             |                         |
|                                                   | 136396 Edgar Tsinanga\>                     | Edgar Tsinanga 136396\> |
+---------------------------------------------------+---------------------------------------------+-------------------------+
| **BBIT 4.2 Group**                                | \<B\>                                       |                         |
+---------------------------------------------------+---------------------------------------------+-------------------------+
| **Course Code**                                   | BBT4206                                     |                         |
+---------------------------------------------------+---------------------------------------------+-------------------------+
| **Course Name**                                   | Business Intelligence II                    |                         |
+---------------------------------------------------+---------------------------------------------+-------------------------+
| **Program**                                       | Bachelor of Business Information Technology |                         |
+---------------------------------------------------+---------------------------------------------+-------------------------+
| **Semester Duration**                             | 21^st^ August 2023 to 28^th^ November 2023  |                         |
+---------------------------------------------------+---------------------------------------------+-------------------------+

# Setup Chunk

**Note:** the following KnitR options have been set as the global defaults:\
`knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE,                        collapse = FALSE, tidy = TRUE)`.

> output:\
> \
> github_document:\
> toc: yes\
> toc_depth: 4\
> fig_width: 6\
> fig_height: 4\
> df_print: default\
> \
> editor_options:\
> chunk_output_type: console

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE,
                      collapse = FALSE, tidy = TRUE)
```

# Loading Dataset

Describe the code chunk here:

```{r Your Second Code Chunk}
student_performance_dataset <- readr::read_csv("data/StudentPerformance.csv",
    col_types = readr::cols(class_group = readr::col_factor(levels = c("A", "B",
        "C")), gender = readr::col_factor(levels = c("1", "0")), YOB = readr::col_date(format = "%Y"),
        regret_choosing_bi = readr::col_factor(levels = c("1", "0")), drop_bi_now = readr::col_factor(levels = c("1",
            "0")), motivator = readr::col_factor(levels = c("1", "0")), read_content_before_lecture = readr::col_factor(levels = c("1",
            "2", "3", "4", "5")), anticipate_test_questions = readr::col_factor(levels = c("1",
            "2", "3", "4", "5")), answer_rhetorical_questions = readr::col_factor(levels = c("1",
            "2", "3", "4", "5")), find_terms_I_do_not_know = readr::col_factor(levels = c("1",
            "2", "3", "4", "5")), copy_new_terms_in_reading_notebook = readr::col_factor(levels = c("1",
            "2", "3", "4", "5")), take_quizzes_and_use_results = readr::col_factor(levels = c("1",
            "2", "3", "4", "5")), reorganise_course_outline = readr::col_factor(levels = c("1",
            "2", "3", "4", "5")), write_down_important_points = readr::col_factor(levels = c("1",
            "2", "3", "4", "5")), space_out_revision = readr::col_factor(levels = c("1",
            "2", "3", "4", "5")), studying_in_study_group = readr::col_factor(levels = c("1",
            "2", "3", "4", "5")), schedule_appointments = readr::col_factor(levels = c("1",
            "2", "3", "4", "5")), goal_oriented = readr::col_factor(levels = c("1",
            "0")), spaced_repetition = readr::col_factor(levels = c("1", "2", "3",
            "4")), testing_and_active_recall = readr::col_factor(levels = c("1",
            "2", "3", "4")), interleaving = readr::col_factor(levels = c("1", "2",
            "3", "4")), categorizing = readr::col_factor(levels = c("1", "2", "3",
            "4")), retrospective_timetable = readr::col_factor(levels = c("1", "2",
            "3", "4")), cornell_notes = readr::col_factor(levels = c("1", "2", "3",
            "4")), sq3r = readr::col_factor(levels = c("1", "2", "3", "4")), commute = readr::col_factor(levels = c("1",
            "2", "3", "4")), study_time = readr::col_factor(levels = c("1", "2",
            "3", "4")), repeats_since_Y1 = readr::col_integer(), paid_tuition = readr::col_factor(levels = c("0",
            "1")), free_tuition = readr::col_factor(levels = c("0", "1")), extra_curricular = readr::col_factor(levels = c("0",
            "1")), sports_extra_curricular = readr::col_factor(levels = c("0", "1")),
        exercise_per_week = readr::col_factor(levels = c("0", "1", "2", "3")), meditate = readr::col_factor(levels = c("0",
            "1", "2", "3")), pray = readr::col_factor(levels = c("0", "1", "2", "3")),
        internet = readr::col_factor(levels = c("0", "1")), laptop = readr::col_factor(levels = c("0",
            "1")), family_relationships = readr::col_factor(levels = c("1", "2",
            "3", "4", "5")), friendships = readr::col_factor(levels = c("1", "2",
            "3", "4", "5")), romantic_relationships = readr::col_factor(levels = c("0",
            "1", "2", "3", "4")), spiritual_wellnes = readr::col_factor(levels = c("1",
            "2", "3", "4", "5")), financial_wellness = readr::col_factor(levels = c("1",
            "2", "3", "4", "5")), health = readr::col_factor(levels = c("1", "2",
            "3", "4", "5")), day_out = readr::col_factor(levels = c("0", "1", "2",
            "3")), night_out = readr::col_factor(levels = c("0", "1", "2", "3")),
        alcohol_or_narcotics = readr::col_factor(levels = c("0", "1", "2", "3")),
        mentor = readr::col_factor(levels = c("0", "1")), mentor_meetings = readr::col_factor(levels = c("0",
            "1", "2", "3")), `Attendance Waiver Granted: 1 = Yes, 0 = No` = readr::col_factor(levels = c("0",
            "1")), GRADE = readr::col_factor(levels = c("A", "B", "C", "D", "E"))),
    locale = readr::locale())
```

# Dimensions

Describe the next code chunk here:

```{r Your Third Code Chunk}
dim(student_performance_dataset)
```

# Data Types

Describe the next code chunk here:

```{r Your Fourth Code Chunk}
sapply(student_performance_dataset, class)

```

# Descriptive Statistics

## Measures of Frequency

Describe the next code chunk here:

```{r Your Fifth Code Chunk}
project_freq <- student_performance_dataset$`Project: Section 5-11: (50%) x/10`
cbind(frequency = table(project_freq),
    percentage = prop.table(table(project_freq)) * 100)
```

## Measures of Central Tendency

Describe the next code chunk here:

```{r Your Sixth Code Chunk}
student_performance_dataset_mode <- names(table(student_performance_dataset$`EXAM: x/60 (60%)`))[
  which(table(student_performance_dataset$`EXAM: x/60 (60%)`) == max(table(student_performance_dataset$`EXAM: x/60 (60%)`)))
]
print(student_performance_dataset_mode)
```
Measures of Distribution/Dispersion/Spread/Scatter/Variability

```{r Your Seventh Code Chunk}
summary(student_performance_dataset)
```
Measure the variance

Describe the next code chunk here:
```{r Your Eighth Code Chunk}
library(dplyr)
library(tidyr)
numeric_columns <- student_performance_dataset %>%
  select_if(is.numeric)

variance_df <- numeric_columns %>%
  summarise_all(~ var(., na.rm = TRUE)) %>%
  pivot_longer(everything(), names_to = "Column", values_to = "Variance")

print(variance_df, n = Inf)
```
Measure the standard deviation

Describe the next code chunk here:
```{r Your Ninth Code Chunk}
standard_deviation <- sd(student_performance_dataset$`CAT 2 (8%): x/100 x 100`, na.rm = TRUE)

print(standard_deviation)
```
Inferential Statistics

Perform ANOVA

Describe the next code chunk here:
```{r Your Tenth Code Chunk}
student_performance_dataset_anova <- aov(`Project: (10%): x/30 x 100 TOTAL` ~ class_group + gender, data = student_performance_dataset)

summary(student_performance_dataset_anova)
```
Qualitative Data Analysis

Univariate Plots

Create Histograms

Describe the next code chunk here:
```{r Your Eleventh Code Chunk}
quiz_score <- as.numeric(student_performance_dataset$`Quizzes and  Bonus Marks (7%): x/79 x 100 TOTAL`)
lab_score <- as.numeric(student_performance_dataset$`Lab Work (7%) x/25 x 100`)
course_score <- as.numeric(student_performance_dataset$`Coursework TOTAL: x/40 (40%)`)

course_score <- course_score[!is.na(course_score)]

par(mfrow = c(2, 2))

# Create a histogram for quiz score
hist(quiz_score,
     main = "Histogram of Project Scores",
     xlab = "Project Scores",
     ylab = "Frequency",
     col = "violet",  
     border = "black",  
     breaks = 10)

# Create a histogram for lab score
hist(lab_score,
     main = "Histogram of Exam Scores",
     xlab = "Exam Scores",
     ylab = "Frequency",
     col = "brown",  
     border = "black",  
     breaks = 10)

# Create a histogram for course scores (after removing missing values)
hist(course_score,
     main = "Histogram of Quiz Scores",
     xlab = "Quiz Scores",
     ylab = "Frequency",
     col = "red",  
     border = "black",  
     breaks = 10)  

```
Missingness Map

Describe the next code chunk here:
```{r Your Twelfth Code Chunk}
if (!is.element("Amelia", installed.packages()[, 1])) {
  install.packages("Amelia", dependencies = TRUE)
}
require("Amelia")

# Create a missing data pattern plot for student_performance_dataset
missmap(student_performance_dataset, col = c("orange", "blue"), legend = TRUE)
```
Multivariate Plots

Create a Scatter Plot

Describe the next code chunk here:
```{r Your Thirteenth Code Chunk}
library(ggplot2)

ggplot(student_performance_dataset, aes(x = gender, y = `A - 1. I am enjoying the subject`)) +
  geom_point() +
  labs(x = "Gender", y = "A  1. I am enjoying the subject") +
  ggtitle("Scatter Plot: Gender vs. I am enjoying the subject")
```

```{r Your Fourteenth Code Chunk}
renv::snapshot()
```

**etc.** as per the lab submission requirements.
